Co-moving stars in \textsl{Gaia DR1}: a catalog of genuine comoving star pairs from low-resolution spectroscopic follow-up

### 2017-06-06 paper/comoving-followup.tex
no most
fix corner.py citation

To derive radial velocities for each source, we then measure the (wavelength)
line centroid of \Ha\ in each corrected spectrum.
We model the absorption line using a pixel-convolved Voigt profile and use a
linear model + a Gaussian process for the background model.
We now use \lbfgsb\ optimization and the resulting maximum-likelihood parameters
to initialize a Markov Chain Monte Carlo (MCMC) sampling of the posterior
probability distribution (PDF) over the 8 parameters: $\ln |A|$, $x_0$,
$\ln\sigma_G$, $\ln\gamma_L$, $\alpha_1$, $\alpha_2$, $\ln\sigma_{3/2}$, and
$\ln\rho_{3/2}$ where the linear background model is defined as $f(\lambda) =
\alpha_1 + \alpha_2\,\lambda$.
The linear amplitude, $A$, is assumed to be negative (i.e. \Ha\ is assumed to be
an absorption feature) and is computed from the corresponding parameter as $A =
-\exp(\ln |A|)$.
We use uniform priors on all parameters except $\ln\rho_{3/2}$, for which we use
a Gaussian; prior parameters are specified in \tablename~\ref{tbl:prior-bounds}.
We use \package{emcee} (\citealt{Foreman-Mackey:2013}) with 64 walkers to
generate samples from the posterior PDF.
We initially run the sampler for 128 steps from the maximum likelihood parameter
values, then resample walker positions in a small ball around the median final
walker positions and run again for 512 steps; together, these two steps
constitute our burn-in phase.
From the final walker positions after burn-in, we run again for 1024 steps and
store these as our posterior samples.
\figurename~\ref{fig:Halpha-mcmc-corner} shows a \package{corner} plot
(\citealt{Foreman-Mackey:2017a}) of all projections of these posterior PDF
samples for a randomly-chosen source (HD\todo{XXXX}, in this case).
The marginal posterior distribution over $x_0$ can be converted to a
distribution over radial velocities; in this case, the estimated (Gaussian
standard deviation) precision is $\approx 2.5~\kms$.
\figurename~\ref{fig:Halpha-mcmc-fit} shows the segment of the source spectrum
around \Ha\ used in the line modeling (black line) with uncertainties (grey
bars).
Top panel shows the linear polynomial + absorption line model computed for the
median sample (dark green line); shaded region shows the 15th to 85th
percentiles computed from all posterior samples.
Bottom panel shows the residuals of the data from the median linear polynomial +
absorption line model (black line) along with the Gaussian process computed from
the median parameters (orange curve) used to model other structure in the
vicinity of \Ha (which can be interpreted as correlated noise).
These figures are produced for all source spectrum fits and visually inspected
to ensure that the MCMC sampling appears to have converged for all cases.

@@

Several of our targets and all of our calibration targets have previously
measured radial velocities (RVs).
We can therefore estimate the absolute velocity uncertainties by comparing our
derived radial velocities with high-quality prior measurements.
Following previous work (\citealt{Shaya:2011}), we search for previous RV
measurements (in decreasing order of priority) with:
\begin{enumerate}
  \item The Geneva-Copenhagen Survey of the Solar Neighborhood
  (\citealt{Nordstrom:2004}).
  \item The Catalogue of Radial Velocities of Galactic Stars with Astrometric
  Data, the Second Version (\citealt{Kharchenko:2007}).
  \item The General Catalogue of Mean Radial Velocities
  (\citealt{Barbier-Brossat:2000}).
  \item The SIMBAD database (\citealt{Wenger:2000}). We use \package{astroquery}
  to retrieve records from the database, which do not contain uncertainties but
  do contain a quality flag. We only retain RVs that have a quality flag set to
  `A' and adopt a $5~\kms$ uncertainty for all such values.
\end{enumerate}
We find that \todo{XXX} of our target stars have reliable prior measurements,
along with the \todo{YYY} calibration targets.
\figurename~\ref{fig:compare-previous} shows the distribution of velocity
differences between our measurements and literature values.
From this distribution, we compute the MAD and estimate the standard deviation
to be $\approx 20~\kms$, which we adopt as our global systematic error for all
absolute radial velocities.
We note, however, that for computing \emph{relative} velocities for pairs of
nearby stars, it is the RV precision that matters most.

@@

\software{
The code used in this project is available from
\url{https://github.com/adrn/GaiaPairsFollowup} under the MIT open-source
software license.
This research utilized the following open-source \python\ packages:
    \package{Astropy} (\citealt{Astropy-Collaboration:2013}),
    \package{astroquery} (\citealt{Ginsburg:2016}),
    \package{ccdproc} (\citealt{Craig:2015}),
    \package{celerite} (\citealt{Foreman-Mackey:2017}),
    \package{emcee} (\citealt{Foreman-Mackey:2013ascl}),
    \package{IPython} (\citealt{Perez:2007}),
    \package{matplotlib} (\citealt{Hunter:2007}),
    \package{numpy} (\citealt{Van-der-Walt:2011}),
    \package{scipy} (\url{https://www.scipy.org/}),
    \package{sqlalchemy} (\url{https://www.sqlalchemy.org/}).
This work additionally used the Gaia science archive
(\url{https://gea.esac.esa.int/archive/}), and the SIMBAD database
(\citealt{Wenger:2000}).
}

>>>>

To derive radial velocities for each source, we then measure the (wavelength)
line centroid of \Ha\ in each corrected spectrum.
We model the absorption line using a pixel-convolved Voigt profile and use a
linear model + a Gaussian process for the background model.
We now use \lbfgsb\ optimization and the resulting maximum-likelihood parameters
to initialize a Markov Chain Monte Carlo (MCMC) sampling of the posterior
probability distribution (PDF) over the 8 parameters: $\ln |A|$, $x_0$,
$\ln\sigma_G$, $\ln\gamma_L$, $\alpha_1$, $\alpha_2$, $\ln\sigma_{3/2}$, and
$\ln\rho_{3/2}$ where the linear background model is defined as $f(\lambda) =
\alpha_1 + \alpha_2\,\lambda$.
The linear amplitude, $A$, is assumed to be negative (i.e. \Ha\ is assumed to be
an absorption feature) and is computed from the corresponding parameter as $A =
-\exp(\ln |A|)$.
We use uniform priors on all parameters except $\ln\rho_{3/2}$, for which we use
a Gaussian; prior parameters are specified in \tablename~\ref{tbl:prior-bounds}.
We use \package{emcee} (\citealt{Foreman-Mackey:2013}) with 64 walkers to
generate samples from the posterior PDF.
We initially run the sampler for 128 steps from the maximum likelihood parameter
values, then resample walker positions in a small ball around the median final
walker positions and run again for 512 steps; together, these two steps
constitute our burn-in phase.
From the final walker positions after burn-in, we run again for 1024 steps and
store these as our posterior samples.
\figurename~\ref{fig:Halpha-mcmc-corner} shows a \package{corner} plot
(\citealt{Foreman-Mackey:2016}) of all projections of these posterior PDF
samples for a randomly-chosen source (HD\todo{XXXX}, in this case).
The marginal posterior distribution over $x_0$ can be converted to a
distribution over radial velocities; in this case, the estimated (Gaussian
standard deviation) precision is $\approx 2.5~\kms$.
\figurename~\ref{fig:Halpha-mcmc-fit} shows the segment of the source spectrum
around \Ha\ used in the line modeling (black line) with uncertainties (grey
bars).
Top panel shows the linear polynomial + absorption line model computed for the
median sample (dark green line); shaded region shows the 15th to 85th
percentiles computed from all posterior samples.
Bottom panel shows the residuals of the data from the median linear polynomial +
absorption line model (black line) along with the Gaussian process computed from
the median parameters (orange curve) used to model other structure in the
vicinity of \Ha (which can be interpreted as correlated noise).
These figures are produced for all source spectrum fits and visually inspected
to ensure that the MCMC sampling appears to have converged for all cases.

@@

Several of our targets and all of our calibration targets have previously
measured radial velocities (RVs).
We can therefore estimate the absolute velocity uncertainties by comparing our
derived radial velocities with high-quality prior measurements.
Following previous work (\citealt{Shaya:2011}), we search for previous RV
measurements (in decreasing order of priority) with:
\begin{enumerate}
  \item The Geneva-Copenhagen Survey of the Solar Neighborhood
  (\citealt{Nordstrom:2004}).
  \item The Catalogue of Radial Velocities of Galactic Stars with Astrometric
  Data, the Second Version (\citealt{Kharchenko:2007}).
  \item The General Catalogue of Mean Radial Velocities
  (\citealt{Barbier-Brossat:2000}).
  \item The SIMBAD database (\citealt{Wenger:2000}). We use \package{astroquery}
  to retrieve records from the database, which do not contain uncertainties but
  do contain a quality flag. We only retain RVs that have a quality flag set to
  `A' and adopt a $5~\kms$ uncertainty for all such values.
\end{enumerate}
We find that \todo{XXX} of our target stars have reliable prior measurements,
along with the \todo{YYY} calibration targets.
\figurename~\ref{fig:compare-previous} shows the distribution of velocity
differences between our measurements and literature values.
From this distribution, we compute the MAD and estimate the standard deviation
to be $\approx 20~\kms$, which we adopt as our global systematic error for all
absolute radial velocities.
We note, however, that for computing \emph{relative} velocities for pairs of
nearby stars, it is the RV precision that matters.

@@

\software{
The code used in this project is available from
\url{https://github.com/adrn/GaiaPairsFollowup} under the MIT open-source
software license.
This research utilized the following open-source \python\ packages:
    \package{Astropy} (\citealt{Astropy-Collaboration:2013}),
    \package{astroquery} (\citealt{Ginsburg:2016}),
    \package{ccdproc} (\citealt{Craig:2015}),
    \package{celerite} (\citealt{Foreman-Mackey:2017}),
    \package{corner} (\citealt{Foreman-Mackey:2016}),
    \package{emcee} (\citealt{Foreman-Mackey:2013ascl}),
    \package{IPython} (\citealt{Perez:2007}),
    \package{matplotlib} (\citealt{Hunter:2007}),
    \package{numpy} (\citealt{Van-der-Walt:2011}),
    \package{scipy} (\url{https://www.scipy.org/}),
    \package{sqlalchemy} (\url{https://www.sqlalchemy.org/}).
This work additionally used the Gaia science archive
(\url{https://gea.esac.esa.int/archive/}), and the SIMBAD database
(\citealt{Wenger:2000}).
}
